Ecological and evolutionary implications of starvation and body size

### 2016-07-21 Manuscript/Yeakel_Kempes-Redner_PNAS_7-19.tex
final small edits before sending to Chris and Sid

\vspace{2mm}
\noindent {\bf Dynamic and energetic barriers to body size} \\ \nonumber
Metabolite transportation constraints are widely thought to place strict boundaries on biological scaling \cite{Brown:1993p708,West:1997cg,Brown:2004wq}, leading to specific predictions on the minimum possible body size for organisms \cite{West:2002ud}.
Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
The \emph{fasting endurance hypothesis} contends that larger body size, being energetically more efficient per unit biomass, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate a selective force towards an optimal upper-bound of ca. $10^7$ grams \cite{Alroy:1998p1594}, the value of which may arise from higher extinction risk for large taxa over evolutionary timescales \cite{Clauset:2009fh}.
These trends are thought to be driven by a combination of climate change and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore this problem.

>>>>

\vspace{2mm}
\noindent {\bf Dynamic and energetic barriers to body size} \\ \nonumber
Metabolite transportation constraints are widely thought to place strict boundaries on biological scaling \cite{Brown:1993p708,West:1997cg,Brown:2004wq}, leading to specific predictions on the minimum possible body size for organisms \cite{West:2002ud}.
Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
The \emph{fasting endurance hypothesis} contends that larger body size, with lower metabolic rates and able to hold more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper-bound of ca. $10^7$ grams \cite{Alroy:1998p1594}, the value of which may arise from higher extinction risk for large taxa over evolutionary timescales \cite{Clauset:2009fh}.
These trends are thought to be driven by a combination of climate change and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore these issues.
